TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-09-20 paper/draft.tex
radii section
mention radius in abstract

\begin{abstract}
\thecannon\ is a flexible, data-driven spectral modeling and parameter inference framework, demonstrated on high-resolution, near-infrared \apogee\ spectra of giant stars to estimate stellar labels (\teff, logg, [Fe/H], and detailed abundances) to precisions higher than the \apogee\ survey model-grid pipelines.
The lack of reliable atmospheric models in the near infrared for temperatures cooler than $\sim3500$\,K motivates the extension of this approach to M-dwarf stars, but with calibrated training labels from optical surveys that overlap \sdssiv\ \apogee.
Using a training sample of 87 M dwarfs in the temperature range $2859 < T_{\rm eff} < 4131$\,K, and metallicty range $-0.48 < \mathrm{[Fe/H]}< 0.49$\,dex, we train a two-parameter model with predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex respectively.
These are parameter-space precision statements, but the model is accurate in the data space: it accurately predicts \apogee\ spectral pixels at the percent level.
We also train a one-dimensional spectral type model using 51 M dwarfs with known spectral types ranging from M0 to M9, with predictive accuracy of 0.9 types. From the \textsl{APOGEE DR14} -- \textsl{Gaia DR2} overlap we select a test sample and employ our trained models to determine temperatures, metallicities and spectral subtypes for nearly all of the M dwarfs in \apogee\ (10,311 sources), and validate our \cannon\ derived \teff\ against color-temperature relations and [Fe/H] against isochrone models in \gaia\ color-magnitude space. Finally we use 
These results have implications for the NASA \tess\ Mission and other projects that involve large samples of M dwarfs.

@@

>>>>

\begin{abstract}
\thecannon\ is a flexible, data-driven spectral modeling and parameter inference framework, demonstrated on high-resolution, near-infrared \apogee\ spectra of giant stars to estimate stellar labels (\teff, logg, [Fe/H], and detailed abundances) to precisions higher than the \apogee\ survey model-grid pipelines.
The lack of reliable atmospheric models in the near infrared for temperatures cooler than $\sim3500$\,K motivates the extension of this approach to M-dwarf stars, but with calibrated training labels from optical surveys that overlap \sdssiv\ \apogee.
Using a training sample of 87 M dwarfs in the temperature range $2859 < T_{\rm eff} < 4131$\,K, and metallicty range $-0.48 < \mathrm{[Fe/H]}< 0.49$\,dex, we train a two-parameter model with predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex respectively.
These are parameter-space precision statements, but the model is accurate in the data space: it accurately predicts \apogee\ spectral pixels at the percent level.
We also train a one-dimensional spectral type model using 51 M dwarfs with known spectral types ranging M0 to M9, to predictive accuracy of 0.9 types. From the \textsl{APOGEE DR14} -- \textsl{Gaia DR2} overlap we select a test sample and employ our trained models to determine temperatures, metallicities and spectral subtypes for nearly all of the M dwarfs in \apogee\ (10,311 sources), and validate our \cannon\ derived \teff\ against color-temperature relations and [Fe/H] against isochrone models in \gaia\ color-magnitude space. Finally we use empirically callibrated relations to estimate radii from $M_K$ absolute magnitudes and \cannon\ temperatures.
These results have implications for the NASA \tess\ Mission and other projects that involve large samples of M dwarfs.

@@

\subsection{Radii}
Using empirically calibrated relations from \citealt{Mann:2015} we determine radii... some diagnostics...
